---
title: "GLM for replicate"
output: html_document
date: "2023-03-23"
---

```{r}
library(ggplot2)
library(dplyr)
library(readr)
library(tidyverse)
library(broom)
```

```{r}
for_colab_RA <- read_csv("https://raw.githubusercontent.com/tranJen/RaBi-Collaboration/main/develop%20model/Data_Akrong.csv")
for_colab_AM <- read_csv("https://raw.githubusercontent.com/tranJen/RaBi-Collaboration/main/develop%20model/Data_McGahan.csv")
```

#### *AM does not have D147E mutation but RA does.*

```{r}
#Manipulate McGahan's data
df_AM <-filter(for_colab_AM[1:428,],library!=0) #remove one has library = 0
df_AM <- filter(df_AM, B1 != 0 | B2 != 0) #remove one has b1=0 and b2=0
df_AM[df_AM==0] <- 0.1

df_AM$B1dMut <- (df_AM$B1/ df_AM$B1[1])  / (df_AM$library/ df_AM$library[1])
df_AM$B2dMut <- (df_AM$B2/ df_AM$B2[1])  / (df_AM$library/ df_AM$library[1])
df_AM$changeB2_B1 <- df_AM$B2dMut - df_AM$B1dMut
```

```{r}
#Manipulate Akrong's data
for_colab_RA <- read_csv("https://raw.githubusercontent.com/tranJen/GLM_Rabi/main/for_colab_RA.csv")
df_RA <-filter(for_colab_RA[1:428,],library!=0) #remove one has library = 0

df_RA$B1 <- (df_RA$B1_1 + df_RA$B1_2 )/2
df_RA$B2 <- (df_RA$B2_1 + df_RA$B2_2 )/2

df_RA <- filter(df_RA, B1 != 0 | B2 != 0) #remove one has b1=0 and b2=0
df_RA[df_RA==0] <- 0.1
df_RA <- subset(df_RA, select = -c(B1_1,B1_2, B2_1, B2_2) )
df_RA$B1dMut <- (df_RA$B1/ df_RA$B1[1])  / (df_RA$library/ df_RA$library[1])
df_RA$B2dMut <- (df_RA$B2/ df_RA$B2[1])  / (df_AM$library/ df_RA$library[1])
df_RA$changeB2_B1 <- df_RA$B2dMut - df_RA$B1dMut

```
# GLM

dMut = (Mut B1/ WT B1) / (Mut lib / WT lib)
```{r}
df <- rbind(df_RA[2:429,], df_AM[2:428, ])
df<- arrange(df, pos)
df$pos <- as.character(df$pos)
```

```{r}
gauss_model <- glm( changeB2_B1 ~ pos * mut , data = df, family = 'gaussian') 
#options(max.print=10000)    #set limit to print the output
with(summary(gauss_model), 1 - deviance/null.deviance)
#The log-likelihood is a measure of the log-probability of the observed data given the model parameters. Higher values of the log-likelihood indicate that the model is a better fit for the data.
logLik(gauss_model)
```
 
"Coefficients: (57 not defined because of singularities)" indicates that 57 of the covariates are a linear combination of some other of the covariates. In other words, it is because of singularity means that the variables are not linearly independent. If you remove the NA variables in the above summary, you will obtain the same result for the rest of the variables. This is because the information given by those variables is already contained in the other variables and thus redundant.

```{r}
fea <- names(coef(gauss_model))[44:length(names(coef(gauss_model)))]      #get features' names

#split the features' names (e.g from "pos137:mutA" to "137" and "A")
mut <- list()
pos <- list ()
for (i in fea){
              str_split <- strsplit(i, ":")
              pos <- append(pos, substr(str_split[[1]][1], 4,6))    # e.g extract 137
              mut <- append(mut, substr(str_split[[1]][2], 4,4))    # e.g extract A
}

est <- coef(gauss_model)[-1][43:length(coef(gauss_model)[-1])]          #get estimated contribution 
#p <- coef(summary(gauss_model))[-1,4]         #get p-value
p <- tidy(gauss_model)$p.value[44:483]        #to extract all including NA value
```

##### *white box is NA values. *

```{r}
#create data frame of pos, mut, estimated contribution, alpha
data_for_plot<- do.call(rbind, Map(data.frame, Pos=pos, Mut=mut, Est = est, p_value =p))
data_for_plot$alpha <- ifelse(data_for_plot$p_value<0.001,'***', ifelse(0.001 <data_for_plot$p_value & data_for_plot$p_value<0.01,'**', ifelse(0.01<data_for_plot$p_value& data_for_plot$p_value<0.05,'*', ifelse(0.05<data_for_plot$p_value& data_for_plot$p_value<0.1,'.', ' '))))
```

```{r}
#plot estimated contribution vs features
ggplot(data_for_plot, aes(x = Pos, y = Mut)) +
  geom_tile(aes(fill = Est)) + 
  geom_text(aes(label = alpha), size = 3)+
  scale_fill_gradient2(midpoint= 0, na.value = 'white',high = 'darkgoldenrod1 ', low = 'cyan4', mid ='aliceblue')+ 
  theme_classic() +
  theme(axis.text.x = element_text(angle = 75, hjust=1), plot.subtitle = element_text(size = 10, color = "red", face = "bold"))+
  labs(y = "Mutation Amino Acid", x = "Enzyme Position", fill= 'Estimated
Contribution', subtitle = 'Signif. codes:  0 *** 0.001 ** 0.01 * 0.05 . 0.1   1')

```
### Remove unsignificant 
```{r}
data_for_plot <- data_for_plot[!(data_for_plot$Pos %in% c(137, 138, 139, 140, 141, 142,144,146,147,150,151,152,153,154,156,158)), ]
```

```{r}
heatmap <- ggplot(data_for_plot, aes(x = Pos, y = Mut)) +
  geom_tile(aes(fill = Est), width = 1, height = 1) +
  geom_text(aes(label = alpha), size = 3) +
  scale_fill_gradient2(midpoint= 0, na.value = 'white', high = 'darkgoldenrod1', low = 'cyan4', mid ='aliceblue') + 
  theme_classic() +
  theme(text =element_text(size = 15),
        # axis.text.x = element_text(angle = 75, hjust = 1), 
        plot.subtitle = element_text(size = 10, color = "red", face = "bold"), 
        aspect.ratio = 3/2,
        # legend.text = element_text(angle = 0, vjust = 1, hjust = 0.5),
        legend.title = element_text(angle = 90, vjust = 0.5, hjust = 0.5, size = 10),
        legend.position =  c(1.2,0.5)) +
  guides(fill = guide_colorbar(title.position = "left", title.hjust = 0.5, label.position = "left"))+
  theme(legend.key.width = unit(0.25, "cm"))+
  labs(y = "Mutation Amino Acid", x = "Enzyme Position", fill= 'Estimated Contribution', subtitle = 'Signif. codes:  0 *** 0.001 ** 0.01 * 0.05 . 0.1   1')
heatmap
ggsave("heatmap_WT.png", heatmap, width = 6, height = 5, units = "in")
```

```{r}
observed <- df$changeB2_B1
fitted <- fitted(gauss_model)
observed <- na.omit(observed)
fitted <- na.omit(fitted)
# Create a data frame for the observed and fitted values
plot_data <- data.frame(Observed = observed, Fitted = fitted)

# Create the scatter plot of observed vs. fitted values using ggplot2
ggplot(plot_data, aes(x = Observed, y = Fitted)) +
  geom_point() +
  xlab("Observed") +
  ylab("Fitted") +
  ggtitle("Observed vs. Fitted") +
  geom_abline(slope = 1, intercept = 0, color = "red")  # Add a 45-degree reference line
```